---
title: "Using Genetic Algorithm with TSP to Solve Resource Distribution During Outlier Event"
author: "Chih-Hao Huang"
output: 
  pdf_document: default
  html_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, results='hide', warning=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(GA)
library(rgdal)
library(tidygeocoder)
```



# 1. Data

*Veterans Health Administration Medical Facilities* data collected from: [link](https://hifld-geoplatform.opendata.arcgis.com/datasets/f11d7d153bfb408f85bd029b2dac9298_0/data)   <br>

VHA data in the US was collected (updated 2018), only contiguous US states including D.C. have been selected. <br>

```{r datasets}
set.seed(2020)
# VHA data
VHA <- read.csv("data/VHA.csv", stringsAsFactors = F)
names(VHA)[
  names(VHA) == "ï..X"
] <- "X"
VHA <- subset(VHA, STATE != "AK")
VHA <- subset(VHA, STATE != "HI")
VHA <- subset(VHA, STATE != "AS")
VHA <- subset(VHA, STATE != "PR")
VHA <- subset(VHA, STATE != "GU")
VHA <- subset(VHA, STATE != "MP")
VHA <- subset(VHA, STATE != "VI")
```


VHA medical centers' rating data collected from *the U.S. Department of Veterans Affairs*: [link](https://www.va.gov/QUALITYOFCARE/measure-up/End_of_Year_Hospital_star_Rating_FY2018.asp) <br>

*The End of Year Hospital Star Rating (FY2018)* contains relative performance star rating of each medical center from 1 to 5 (5 being the best). <br>

```{r}
VHAr <- read.csv("data/VHArating.csv", stringsAsFactors = F)
VHAr <- VHAr[,1:3]
colnames(VHAr) <- c("VSIN", "NAME", "Performance")
```

Subset facilities name, longitude and latitude from `VHA`. Then select VHA medical centers' only from `VHA`, and merge with `VHAr`

```{r}
mc <- subset(VHA, select = c("NAME", "LONGITUDE", "LATITUDE", "STATE"))
mc <- subset(mc, NAME %in% VHAr$NAME)
# Remove duplicated data
mc <- mc[!duplicated(mc$NAME),]
VHAr <- VHAr[!duplicated(VHAr$NAME),]
# Order A->Z
mc <- mc[order(mc$NAME),]
VHAr <- VHAr[order(VHAr$NAME),]
# Bind Columns
mc <- cbind(mc, VHAr)
mc <- subset(mc, select = c(1:5, 7))
```

The *Provisional COVID-19 Death Counts by Sex, Age, and State* was collected from *CDC*: [link](https://data.cdc.gov/NCHS/Provisional-COVID-19-Death-Counts-by-Sex-Age-and-S/9bhg-hcku) <br>

In this study, only contiguous states and D.C. are selected. <br>

```{r}
# CDC COVID-19 Death Count Dataset
covid <- read.csv("data/CDCdeathCOVID.csv", stringsAsFactors = F)
# COVID-related specifically
covid <- subset(covid, select = -c(1:3,13))
# Keep data for each states, each sex, and each age group only
covid <- subset(covid, State != "United States" & Sex != "All" & Age.group != "All ages")
# Remove data for non-contiguous state (excluding D.C)
covid <- subset(covid, State != "Puerto Rico")
covid <- subset(covid, State != "Alaska")
covid <- subset(covid, State != "Hawaii")
covid <- subset(covid, State != "New York City")
# Replace NA with 0
covid[is.na(covid)] <- 0

# Replace State name with Abbreviation
covid$State <- state.abb[match(covid$State, state.name)]
covid$State[is.na(covid$State)] <- "DC"
```

Select COVID only patients for now
```{r}
cv <- covid[,1:4]
```


State center geological coordinates of the contiguous states and D.C. were collected from `state` function. <br>

```{r}
# Collect State center information
sc <- data.frame(
  matrix(
    0,
    nrow = 49,
    ncol = 2
  )
)
sc <- data.frame(
  State = state.abb,
  long = state.center$x,
  lat = state.center$y
)
sc <- subset(sc, State != "AK")
sc <- subset(sc, State != "HI")
dc <- data.frame(
  State = "DC",
  long = -77.0369,
  lat = 38.9072
)
sc <- bind_rows(sc, dc)
sc <- sc[order(sc$State),]
```


Shape files of the US collected from: [link](https://www.census.gov/cgi-bin/geo/shapefiles/index.php)   <br>

The shape files of the contiguous US states including D.C. used are based on data collected in 2019. <br>

```{r}
# Contiguous US (including D.C.) map shape
US <- readOGR(dsn = "map_shapes", layer = "tl_2019_us_state")
US <- subset(US, STUSPS != "AK")
US <- subset(US, STUSPS != "HI")
US <- subset(US, STUSPS != "AS")
US <- subset(US, STUSPS != "PR")
US <- subset(US, STUSPS != "GU")
US <- subset(US, STUSPS != "MP")
US <- subset(US, STUSPS != "VI")
US <- fortify(US)
```


Voinsky, Irena et al. “Effects of age and sex on recovery from COVID-19: Analysis of 5769 Israeli patients.” *The Journal of infection* vol. 81,2 (2020): e102-e103. doi:10.1016/j.jinf.2020.05.026 <br> [link](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7229949/#) <br>

*Note: This data is based on Israeli patients, for more accurate result, data on US patient should be used.* <br>

```{r}
# Age group
ag <- c("19 and below", "20-29", "30-39", "40-49", "50-59", "60 and above")

# Age group in Male
rm <- c(13.61, 13.97, 14.46, 14.79, 14.81, 14.73)

# Age group in Female
rf <- c(13.24, 13.99, 14.17, 14.76, 14.18, 13.99)
```

Alter age group making data frames compatible <br>
```{r}
ag <- cv$Age.group[1:11]
rm <- c(rm[1], rm[1], rm[1], mean(rm[1], rm[2]), mean(rm[2], rm[3]), mean(rm[3], rm[4]), mean(rm[4], rm[5]), mean(rm[5], rm[6]), rm[6], rm[6], rm[6])

rf <- c(rf[1], rf[1], rf[1], mean(rf[1], rf[2]), mean(rf[2], rf[3]), mean(rf[3], rf[4]), mean(rf[4], rf[5]), mean(rf[5], rf[6]), rf[6], rf[6], rf[6])

# Create data frame 
rp <- data.frame(
  matrix(c(rm, rf), nrow = 2 * length(rm), ncol = 1)
)

sex <- data.frame(
  matrix(c(
    rep("Male", length(rm)), rep("Female", length(rf))
  ), 
    nrow = 2 * length(rm), 
    ncol = 1
  )
)

ag <- data.frame(
  matrix(
    c(ag, ag),
    nrow = 2* length(rm),
    ncol = 1
  )
)

rp <- bind_cols(sex, ag, rp)

colnames(rp) <- c("Sex", "Age", "RecoveryPeriod")
```

According to America's Health Insurance Plans, the average cost per patient per day is around $1500. <br>

```{r}
ccpd <- 1500
```

Therefore, the average cost per person for COVID treatment in each group will be the following: <br>

```{r}
rp$cost <- rp$RecoveryPeriod * ccpd
```

Calculate cost of treatment in each states
```{r}
# Calculate cost of treatment based on sex, age group and states.
cv$cost <- ifelse(cv$Sex == rp$Sex & cv$Age.group == rp$Age, cv$COVID.19.Deaths * rp$cost)

# Total cost of treatment in each states
covidState <- aggregate(cbind(cost) ~ State, data = cv, FUN = sum)
# Normalized cost data into index between 0 and 1
covidState$costID <- (covidState$cost - min(covidState$cost)) / (max(covidState$cost) - min(covidState$cost))
```







# 2. Functions
Following functions are based on: [link](https://rpubs.com/somasdhavala/GAeg). This function calculate the total distance based on given tour (TSP): <br>
```{r}
tourLength <- function(tour, distMatrix) {
    tour <- c(tour, tour[1])
    route <- embed(tour, 2)[, 2:1]
    sum(distMatrix[route])
}
```

In this fitness function, only distance have been considered. <br>
```{r}
fitness1 <- function(tour, ...) 1/tourLength(tour, ...)
```








# 3. State Centers without Route
Distance matrix between states centers <br>
```{r}
scd <- as.matrix(stats::dist(select(sc, long, lat), diag = TRUE, upper = TRUE))
colnames(scd) <- sc$State
rownames(scd) <- sc$State
```

```{r}
p1 <- ggplot() +
  geom_polygon(
    data = US,
    aes(
      x = long,
      y = lat,
      group = group,
      fill = NA
    ),
    color = "grey35",
    fill = NA,
    size = 0.5
  ) + geom_point(
    data = sc,
    shape = 13,
    size = 5,
    color = "seagreen",
    aes(
      x = long,
      y = lat,
      
    )
  ) + coord_map() + labs(
  x = NULL,
  y = NULL,
  title = "US States' Center"
) + theme(
  axis.line =  element_blank(),
  axis.text =  element_blank(),
  axis.ticks =  element_blank(),
  legend.position = "none",
  plot.title = element_text(hjust = 0.5, size = 18),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank()
)

p1

# Save image file
pdf(
  file = "image/p1.pdf",
  width = 12,
  height = 7
)

p1

dev.off()

```






# 4. VHA without Route
Distance matrix between medical centers <br>
```{r}
distance <- as.matrix(stats::dist(select(mc, LONGITUDE, LATITUDE), diag = TRUE, upper = TRUE))
colnames(distance) <- mc$NAME
rownames(distance) <- mc$NAME
```

Veterans Health Administration Medical Facilities data on US map. Size based on the Performance Rating from 1 to 5. Color based on the state.
```{r}
p2 <- ggplot() +
  geom_polygon(
    data = US,
    aes(
      x = long,
      y = lat,
      group = group,
      fill = NA
    ),
    color = "grey35",
    fill = NA,
    size = 0.5
  ) + geom_point(
    data = mc,
    alpha = 0.6,
    aes(
      x = LONGITUDE,
      y = LATITUDE,
      size = Performance,
      color = STATE
    )
  ) + coord_map() + labs(
  x = NULL,
  y = NULL,
  title = "VHA Medical Center"
) + theme(
  axis.line =  element_blank(),
  axis.text =  element_blank(),
  axis.ticks =  element_blank(),
  legend.position = "none",
  plot.title = element_text(hjust = 0.5, size = 18),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank()
)

p2

# Save image file
pdf(
  file = "image/p2.pdf",
  width = 15,
  height = 7
)

p2

dev.off()

```





# 5. VHA and US State Center
```{r}
p3 <- ggplot() +
  geom_polygon(
    data = US,
    aes(
      x = long,
      y = lat,
      group = group,
      fill = NA
    ),
    color = "grey35",
    fill = NA,
    size = 0.5
  ) + geom_point(
    data = sc,
    shape = 13,
    size = 5,
    aes(
      x = long,
      y = lat,
      color = State
    )
  ) + geom_point(
    data = mc,
    alpha = 0.3,
    aes(
      x = LONGITUDE,
      y = LATITUDE,
      color = STATE,
      size = Performance
    )
  ) + coord_map() + labs(
  x = NULL,
  y = NULL,
  title = "U.S. State Centers with VHA Medical Centers"
) + theme(
  axis.line =  element_blank(),
  axis.text =  element_blank(),
  axis.ticks =  element_blank(),
  legend.position = "none",
  plot.title = element_text(hjust = 0.5, size = 18),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank()
)

p3

# Save image file
pdf(
  file = "image/p3.pdf",
  width = 12,
  height = 7
)

p3

dev.off()
```







# 6. Suggested Route Between States' Center Based on GA and TSP (Distance Only)

```{r}
nr <- nrow(sc)
GA.fit <- ga(
  type = "permutation", 
  fitness = fitness1, 
  distMatrix = scd, 
  lower = 1, 
  upper = nr, 
  popSize = 100, 
  maxiter = 5000, 
  run = 1000, 
  pmutation = 0.2, 
  monitor = NULL
)

summary(GA.fit)
```

```{r}
n = nrow(sc)
id = list(1:n)
id <- as.data.frame(id)
tour <- GA.fit@solution[1,]
tour <- as.data.frame(tour)
path <- list(1:n)
path <- as.data.frame(path)
tour <- cbind(tour, path)
colnames(tour) <- c("id", "path_order")
vis <- cbind(id, sc)
colnames(vis)[1] <- "id"
vis <- merge(vis, tour)
vis <- vis %>% 
  arrange(path_order)

p4 <- ggplot() +
  geom_polygon(
    data = US,
    aes(
      x = long,
      y = lat,
      group = group,
      fill = NA
    ),
    color = "grey35",
    fill = NA,
    size = 0.5
  ) + geom_point(
    data = sc,
    size = 5,
    shape = 13,
    aes(
      x = long,
      y = lat
    )
  ) + coord_map() + labs(
  x = NULL,
  y = NULL,
  title = "State Centers – Trad. TSP - Unnormalized"
) + geom_path(
  data = vis, 
  size = 1.2,
  aes(
    x = long,
    y = lat,
    alpha = 0.9
  ),
  color = "green4",
) + theme(
  axis.line =  element_blank(),
  axis.text =  element_blank(),
  axis.ticks =  element_blank(),
  legend.position = "none",
  plot.title = element_text(hjust = 0.5, size = 18),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank()
)

p4

# Save image file
pdf(
  file = "image/p4.pdf",
  width = 12,
  height = 7
)

p4

dev.off()

jpeg(
  file = "image/sc_ff1_un.jpeg"
)

p4

dev.off()
```










# 7. Suggested Route Between VHA Medical Centers Based on GA and TSP (Distance Only)
```{r}
nr <- nrow(mc)
GA.fit <- ga(
  type = "permutation", 
  fitness = fitness1, 
  distMatrix = distance, 
  lower = 1, 
  upper = nr, 
  popSize = 100, 
  maxiter = 5000, 
  run = 1000, 
  pmutation = 0.2, 
  monitor = NULL
)

summary(GA.fit)
```

```{r}
n = nrow(mc)
id = list(1:n)
id <- as.data.frame(id)
tour <- GA.fit@solution[1,]
tour <- as.data.frame(tour)
path <- list(1:n)
path <- as.data.frame(path)
tour <- cbind(tour, path)
colnames(tour) <- c("id", "path_order")
vis <- cbind(id, mc)
colnames(vis)[1] <- "id"
vis <- merge(vis, tour)
vis <- vis %>% 
  arrange(path_order)

p5 <- ggplot() +
  geom_polygon(
    data = US,
    aes(
      x = long,
      y = lat,
      group = group,
      fill = NA
    ),
    color = "grey35",
    fill = NA,
    size = 0.5
  ) + geom_point(
    data = mc,
    aes(
      x = LONGITUDE,
      y = LATITUDE,
      size = Performance,
      color = STATE,
    )
  ) + coord_map() + labs(
  x = NULL,
  y = NULL,
  title = "VAH - Trad. TSP - Unnormalized"
) + geom_path(
  data = vis,
  size = 1.2,
  aes(
    x = LONGITUDE,
    y = LATITUDE,
    alpha = 0.9
  ),
  color = "orange4",
) + theme(
  axis.line =  element_blank(),
  axis.text =  element_blank(),
  axis.ticks =  element_blank(),
  legend.position = "none",
  plot.title = element_text(hjust = 0.5, size = 18),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank()
)

p5

# Save image file
pdf(
  file = "image/p5.pdf",
  width = 12,
  height = 7
)

p5

dev.off()

jpeg(
  file = "image/vah_ff1_un.jpeg"
)

p5

dev.off()
```




# 8. GA using Patients, Hospitals, and TSP
Here I assume *the state center is the distribution center for each state*.

In the data, we have the number of patients based on their states, sex, and age group. These information were used to calculate the *total cost* to treat them based on number of patients in each state, the cost per day for the patient treatment and the recovery period data based on sex and age group. <br>

`Cost` is based on `Sex`, and `Age`. Different `Sex` and `Age` group have different average `Recovery Period`, therefore the cost will be different. `Cost` and `Recovery Period` are proportional in this study. Here average cost per day per COVID-19 patient was used. <br>

```{r}
# Total patients per state
sc$cost <- covidState$cost
patient <- aggregate(cbind(COVID.19.Deaths) ~ State, data = covid, FUN = sum)
sc$patient <- patient$COVID.19.Deaths
```

`Hospital performance rating` include hospital staffs and resources and other factors. Here this will be use to calculate index which shows the total resources in each state. <br>
```{r}
# Total hospital performance per state
hp <- aggregate(cbind(Performance) ~ STATE, data = mc, FUN = sum)
sc$hp <- hp$Performance
```

In the data, there are two states without patient data, these will be removed from this study. <br>
```{r}
# Remove states with no patients
sc <- subset(sc, patient != 0)
```

Calculate hospital resources per patient index using:
\[
  \frac{Hospital Performance Rating}{Patient}
\]
```{r}
sc$hrp <- sc$hp / sc$patient
```

The lower the above index, less resources per patient in the state, which will be more emergeny then other states. <br>

## GA
Now apply the data to the GA fitness function based on the following function: <br>
\[
    \frac{Patient}{Cost \times Distance \times Hospital Performance Rating} 
\]

```{r}
scd <- as.matrix(stats::dist(select(sc, long, lat), diag = TRUE, upper = TRUE))
scd <- as.data.frame(scd)
colnames(scd) <- sc$State
rownames(scd) <- sc$State
```

```{r}
sc$crp <- sc$cost * sc$hrp
crp <- sc$crp
crp <- as.data.frame(crp)
row.names(crp) <- sc$State

# transpose
crp <- t(crp)
crp <- as.data.frame(crp)

# Create new distance matrix accounting all variables
scd_w <- data.frame(mapply(`*`, scd, crp))
colnames(scd_w) <- sc$State
row.names(scd_w) <- sc$State
```

Apply the fitness function to GA to find out the best route.
```{r}
nr <- nrow(scd_w)
GA.fit <- ga(
  type = "permutation", 
  fitness = fitness1, 
  distMatrix = scd_w, 
  lower = 1, 
  upper = nr, 
  popSize = 100, 
  maxiter = 5000, 
  run = 1000, 
  pmutation = 0.2, 
  monitor = NULL
)

summary(GA.fit)
```


```{r}
n = nrow(sc)
id = list(1:n)
id <- as.data.frame(id)
tour <- GA.fit@solution[1,]
tour <- as.data.frame(tour)
path <- list(1:n)
path <- as.data.frame(path)
tour <- cbind(tour, path)
colnames(tour) <- c("id", "path_order")
vis <- cbind(id, sc)
colnames(vis)[1] <- "id"
vis <- merge(vis, tour)
vis <- vis %>% 
  arrange(path_order)

p6 <- ggplot() +
  geom_polygon(
    data = US,
    aes(
      x = long,
      y = lat,
      group = group,
      fill = NA
    ),
    color = "grey35",
    fill = NA,
    size = 0.5
  ) + geom_point(
    data = sc,
    shape = 13,
    aes(
      x = long,
      y = lat,
      size = (1/crp)
    )
  ) + coord_map() + labs(
  x = NULL,
  y = NULL,
  title = "State Centers - FF2 - Unnormalzied"
) + geom_path(
  data = vis,
  size = 1.2,
  aes(
    x = long,
    y = lat,
    alpha = 0.9
  ),
  color = "darkorchid4",
) + theme(
  axis.line =  element_blank(),
  axis.text =  element_blank(),
  axis.ticks =  element_blank(),
  legend.position = "none",
  plot.title = element_text(hjust = 0.5, size = 18),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank()
)

p6

# Save image file
pdf(
  file = "image/p6.pdf",
  width = 12,
  height = 7
)

p6

dev.off()


jpeg(
  file = "image/sc_ff2_un.jpeg"
  
)

p6

dev.off()
```













# 8. K-means Clustering

Cluster state center using k-means clustering approach into *four* clusters. <br> 

*Note: Number of clusters and cluster centers can be changed based on different situations, such as amount of distribution center and location of the distribution center.* <br>

```{r}
# Clustering based on distance
km <- kmeans(scd, centers = 4)

sc$region <- km$cluster
```


```{r}
p7 <- ggplot() +
  geom_polygon(
    data = US,
    aes(
      x = long,
      y = lat,
      group = group,
      fill = NA
    ),
    color = "grey35",
    fill = NA,
    size = 0.5
  ) + geom_point(
    data = sc,
    shape = 13,
    size = 5,
    aes(
      x = long,
      y = lat,
      colour = factor(region),
      alpha = 0.9
    )
  ) + coord_map() +
  scale_fill_brewer(
    palette = "set1"
  )+ labs(
  x = NULL,
  y = NULL,
  title = "K-means State Clusters"
) + theme(
  axis.line =  element_blank(),
  axis.text =  element_blank(),
  axis.ticks =  element_blank(),
  legend.position = "none",
  plot.title = element_text(hjust = 0.5, size = 18),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank()
)

p7

# Save image file
pdf(
  file = "image/p7.pdf",
  width = 12,
  height = 7
)

p7

dev.off()
```

Subset dataframe into four sub dataframe based on region cluster: <br>

```{r}
# Region 1
sc1 <- subset(sc, region == 1)
sccoord1 <- subset(sc1, select = c("long", "lat"))
row.names(sccoord1) <- sc1$State

####################################################

# Region 2
sc2 <- subset(sc, region == 2)
sccoord2 <- subset(sc2, select = c("long", "lat"))
row.names(sccoord2) <- sc2$State

####################################################

# Region 3
sc3 <- subset(sc, region == 3)
sccoord3 <- subset(sc3, select = c("long", "lat"))
row.names(sccoord3) <- sc3$State

####################################################

# Region 4
sc4 <- subset(sc, region == 4)
sccoord4 <- subset(sc4, select = c("long", "lat"))
row.names(sccoord4) <- sc4$State
```


Distance matrix for each region
```{r}
# Region 1
scd1 <- as.matrix(stats::dist(select(sc1, long, lat), diag = TRUE, upper = TRUE))
scd1 <- as.data.frame(scd1)
colnames(scd1) <- sc1$State
rownames(scd1) <- sc1$State

####################################################

# Region 2
scd2 <- as.matrix(stats::dist(select(sc2, long, lat), diag = TRUE, upper = TRUE))
scd2 <- as.data.frame(scd2)
colnames(scd2) <- sc2$State
rownames(scd2) <- sc2$State

####################################################

# Region 3
scd3 <- as.matrix(stats::dist(select(sc3, long, lat), diag = TRUE, upper = TRUE))
scd3 <- as.data.frame(scd3)
colnames(scd3) <- sc3$State
rownames(scd3) <- sc3$State

####################################################

# Region 4
scd4 <- as.matrix(stats::dist(select(sc4, long, lat), diag = TRUE, upper = TRUE))
scd4 <- as.data.frame(scd4)
colnames(scd4) <- sc4$State
rownames(scd4) <- sc4$State
```




Create new distance matrix for each region for fitness function based on:
\[
    \frac{Patient}{Cost \times Distance \times Hospital Performance Rating} 
\]

```{r}
# Region 1
sc1$crp <- sc1$cost * sc1$hrp
crp1 <- sc1$crp
crp1 <- as.data.frame(crp1)
row.names(crp1) <- sc1$State

## transpose
crp1 <- t(crp1)
crp1 <- as.data.frame(crp1)

## Create new distance matrix accounting all variables
scd1_w <- data.frame(mapply(`*`, scd1, crp1))
colnames(scd1_w) <- sc1$State
row.names(scd1_w) <- sc1$State

####################################################

# Region 2
sc2$crp <- sc2$cost * sc2$hrp
crp2 <- sc2$crp
crp2 <- as.data.frame(crp2)
row.names(crp2) <- sc2$State

## transpose
crp2 <- t(crp2)
crp2 <- as.data.frame(crp2)

## Create new distance matrix accounting all variables
scd2_w <- data.frame(mapply(`*`, scd2, crp2))
colnames(scd2_w) <- sc2$State
row.names(scd2_w) <- sc2$State

####################################################

# Region 3
sc3$crp <- sc3$cost * sc3$hrp
crp3 <- sc3$crp
crp3 <- as.data.frame(crp3)
row.names(crp3) <- sc3$State

## transpose
crp3 <- t(crp3)
crp3 <- as.data.frame(crp3)

## Create new distance matrix accounting all variables
scd3_w <- data.frame(mapply(`*`, scd3, crp3))
colnames(scd3_w) <- sc3$State
row.names(scd3_w) <- sc3$State

####################################################

# Region 4
sc4$crp <- sc4$cost * sc4$hrp
crp4 <- sc4$crp
crp4 <- as.data.frame(crp4)
row.names(crp4) <- sc4$State

## transpose
crp4 <- t(crp4)
crp4 <- as.data.frame(crp4)

## Create new distance matrix accounting all variables
scd4_w <- data.frame(mapply(`*`, scd4, crp4))
colnames(scd4_w) <- sc4$State
row.names(scd4_w) <- sc4$State
```

Apply the fitness function to GA for each region.

```{r}
# Region 1
nr <- nrow(scd1_w)
GA.fitr1 <- ga(
  type = "permutation", 
  fitness = fitness1, 
  distMatrix = scd1_w, 
  lower = 1, 
  upper = nr, 
  popSize = 100, 
  maxiter = 5000, 
  run = 1000, 
  pmutation = 0.2, 
  monitor = NULL
)

summary(GA.fitr1)

####################################################

# Region 2
nr <- nrow(scd2_w)
GA.fitr2 <- ga(
  type = "permutation", 
  fitness = fitness1, 
  distMatrix = scd2_w, 
  lower = 1, 
  upper = nr, 
  popSize = 100, 
  maxiter = 5000, 
  run = 1000, 
  pmutation = 0.2, 
  monitor = NULL
)

summary(GA.fitr2)

####################################################

# Region 3
nr <- nrow(scd3_w)
GA.fitr3 <- ga(
  type = "permutation", 
  fitness = fitness1, 
  distMatrix = scd3_w, 
  lower = 1, 
  upper = nr, 
  popSize = 100, 
  maxiter = 5000, 
  run = 1000, 
  pmutation = 0.2, 
  monitor = NULL
)

summary(GA.fitr3)

####################################################

# Region 4
nr <- nrow(scd4_w)
GA.fitr4 <- ga(
  type = "permutation", 
  fitness = fitness1, 
  distMatrix = scd4_w, 
  lower = 1, 
  upper = nr, 
  popSize = 100, 
  maxiter = 5000, 
  run = 1000, 
  pmutation = 0.2, 
  monitor = NULL
)

summary(GA.fitr4)
```


Prepare data for visuallization

```{r}
# Region 1
n = nrow(sc1)
id = list(1:n)
id <- as.data.frame(id)
tour <- GA.fitr1@solution[1,]
tour <- as.data.frame(tour)
path <- list(1:n)
path <- as.data.frame(path)
tour <- cbind(tour, path)
colnames(tour) <- c("id", "path_order")
vis1 <- cbind(id, sc1)
colnames(vis1)[1] <- "id"
vis1 <- merge(vis1, tour)
vis1 <- vis1 %>% 
  arrange(path_order)

####################################################

# Region 2
n = nrow(sc2)
id = list(1:n)
id <- as.data.frame(id)
tour <- GA.fitr2@solution[1,]
tour <- as.data.frame(tour)
path <- list(1:n)
path <- as.data.frame(path)
tour <- cbind(tour, path)
colnames(tour) <- c("id", "path_order")
vis2 <- cbind(id, sc2)
colnames(vis2)[1] <- "id"
vis2 <- merge(vis2, tour)
vis2 <- vis2 %>% 
  arrange(path_order)

####################################################

# Region 3
n = nrow(sc3)
id = list(1:n)
id <- as.data.frame(id)
tour <- GA.fitr3@solution[1,]
tour <- as.data.frame(tour)
path <- list(1:n)
path <- as.data.frame(path)
tour <- cbind(tour, path)
colnames(tour) <- c("id", "path_order")
vis3 <- cbind(id, sc3)
colnames(vis3)[1] <- "id"
vis3 <- merge(vis3, tour)
vis3 <- vis3 %>% 
  arrange(path_order)

####################################################

# Region 4
n = nrow(sc4)
id = list(1:n)
id <- as.data.frame(id)
tour <- GA.fitr4@solution[1,]
tour <- as.data.frame(tour)
path <- list(1:n)
path <- as.data.frame(path)
tour <- cbind(tour, path)
colnames(tour) <- c("id", "path_order")
vis4 <- cbind(id, sc4)
colnames(vis4)[1] <- "id"
vis4 <- merge(vis4, tour)
vis4 <- vis4 %>% 
  arrange(path_order)
```



Visualization


```{r}
p8 <- ggplot() +
  geom_polygon(
    data = US,
    aes(
      x = long,
      y = lat,
      group = group,
      fill = NA
    ),
    color = "grey35",
    fill = NA,
    size = 0.5
  ) + geom_point(
    data = sc,
    shape = 13,
    aes(
      x = long,
      y = lat,
      colour = factor(region),
      size = (1/crp),
    )
  ) + geom_path(
  data = vis1, 
  color = "red",
  size = 1.2,
  aes(
    x = long,
    y = lat,
    alpha = 0.9
  )
)+ geom_path(
  data = vis2,
  color = "green",
  size = 1.2,
  aes(
    x = long,
    y = lat,
    alpha = 0.9
  )
)+ geom_path(
  data = vis3, 
  color = "blue",
  size = 1.2,
  aes(
    x = long,
    y = lat,
    alpha = 0.9
  )
)+ geom_path(
  data = vis4, 
  color = "purple",
  size = 1.2,
  aes(
    x = long,
    y = lat,
    alpha = 0.9
  )
) + coord_map() + labs(
  x = NULL,
  y = NULL,
  title = "SC - Kmeans - FF2 - Unnormalized"
) + theme(
  axis.line =  element_blank(),
  axis.text =  element_blank(),
  axis.ticks =  element_blank(),
  legend.position = "none",
  plot.title = element_text(hjust = 0.5, size = 18),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank()
)

p8

# Save image file
pdf(
  file = "image/p8.pdf",
  width = 12,
  height = 7
)

p8

dev.off()

jpeg(
  file = "image/sc_ff2_un_kmeans.jpeg"
)

p8

dev.off()
```





# 9. Using Normalized Values


```{r}
scd <- as.matrix(stats::dist(select(sc, long, lat), diag = TRUE, upper = TRUE))
scd <- as.data.frame(scd)
colnames(scd) <- sc$State
rownames(scd) <- sc$State
scd_n <- scd/max(scd)
```

```{r}
sc_n <- sc
sc_n$cost <- sc_n$cost/max(sc_n$cost)
sc_n$patient <- sc_n$patient/max(sc_n$patient)
sc_n$hp <- sc_n$hp/max(sc_n$hp)
sc_n$hrp <- sc_n$hp/sc_n$patient
sc_n$hrp <- sc_n$hrp/max(sc_n$hrp)
sc_n$crp <- sc_n$cost * sc_n$hrp
sc_n$crp <- sc_n$crp/max(sc_n$crp)

crp_n <- sc_n$crp
crp_n <- as.data.frame(crp_n)


# transpose
crp_n <- t(crp_n)
crp_n <- as.data.frame(crp_n)

# Create new distance matrix accounting all variables
scd_nw <- data.frame(mapply(`*`, scd_n, crp_n))
colnames(scd_nw) <- sc_n$State
row.names(scd_nw) <- sc_n$State
```

Apply the fitness function to GA to find out the best route.
```{r}
nr <- nrow(scd_nw)
GA.fit <- ga(
  type = "permutation", 
  fitness = fitness1, 
  distMatrix = scd_nw, 
  lower = 1, 
  upper = nr, 
  popSize = 100, 
  maxiter = 5000, 
  run = 1000, 
  pmutation = 0.2, 
  monitor = NULL
)

summary(GA.fit)
```

Visualization
```{r}
n = nrow(sc_n)
id = list(1:n)
id <- as.data.frame(id)
tour <- GA.fit@solution[1,]
tour <- as.data.frame(tour)
path <- list(1:n)
path <- as.data.frame(path)
tour <- cbind(tour, path)
colnames(tour) <- c("id", "path_order")
vis <- cbind(id, sc_n)
colnames(vis)[1] <- "id"
vis <- merge(vis, tour)
vis <- vis %>% 
  arrange(path_order)

p9 <- ggplot() +
  geom_polygon(
    data = US,
    aes(
      x = long,
      y = lat,
      group = group,
      fill = NA
    ),
    color = "grey35",
    fill = NA,
    size = 0.5
  ) + geom_point(
    data = sc_n,
    shape = 13,
    aes(
      x = long,
      y = lat,
      size = (1/crp)
    )
  ) + coord_map() + labs(
  x = NULL,
  y = NULL,
  title = "State Centers - FF2 - Normalized"
) + geom_path(
  data = vis,
  size = 1.2,
  aes(
    x = long,
    y = lat,
    alpha = 0.9
  ),
  color = "firebrick1",
) + theme(
  axis.line =  element_blank(),
  axis.text =  element_blank(),
  axis.ticks =  element_blank(),
  legend.position = "none",
  plot.title = element_text(hjust = 0.5, size = 18),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank()
)

p9

# Save image file
pdf(
  file = "image/p9.pdf",
  width = 12,
  height = 7
)

p9

dev.off()

jpeg(
  file = "image/sc_ff2_nor.jpeg"
)

p9

dev.off()
```






# 10. K-means Clustering with Normalized Data

Cluster state center using k-means clustering approach into *four* clusters. <br> 

*Note: Number of clusters and cluster centers can be changed based on different situations, such as amount of distribution center and location of the distribution center.* <br>

```{r}
# Clustering based on distance
km_n <- kmeans(scd_n, centers = 4)

sc_n$region <- km_n$cluster
```


Subset dataframe into four sub dataframe based on region cluster: <br>

```{r}
# Region 1
sc1_n <- subset(sc_n, region == 1)
sccoord1_n <- subset(sc1_n, select = c("long", "lat"))
row.names(sccoord1_n) <- sc1_n$State

####################################################

# Region 2
sc2_n <- subset(sc_n, region == 2)
sccoord2_n <- subset(sc2_n, select = c("long", "lat"))
row.names(sccoord2_n) <- sc2_n$State

####################################################

# Region 3
sc3_n <- subset(sc_n, region == 3)
sccoord3_n <- subset(sc3_n, select = c("long", "lat"))
row.names(sccoord3_n) <- sc3_n$State

####################################################

# Region 4
sc4_n <- subset(sc_n, region == 4)
sccoord4_n <- subset(sc4_n, select = c("long", "lat"))
row.names(sccoord4_n) <- sc4_n$State
```


Distance matrix for each region
```{r}
# Region 1
scd1_n <- as.matrix(stats::dist(select(sc1_n, long, lat), diag = TRUE, upper = TRUE))
scd1_n <- as.data.frame(scd1_n)
colnames(scd1_n) <- sc1_n$State
rownames(scd1_n) <- sc1_n$State
scd1_n <- scd1_n/max(scd1_n)

####################################################

# Region 2
scd2_n <- as.matrix(stats::dist(select(sc2_n, long, lat), diag = TRUE, upper = TRUE))
scd2_n <- as.data.frame(scd2_n)
colnames(scd2_n) <- sc2_n$State
rownames(scd2_n) <- sc2_n$State
scd2_n <- scd2_n/max(scd2_n)

####################################################

# Region 3
scd3_n <- as.matrix(stats::dist(select(sc3_n, long, lat), diag = TRUE, upper = TRUE))
scd3_n <- as.data.frame(scd3_n)
colnames(scd3_n) <- sc3_n$State
rownames(scd3_n) <- sc3_n$State
scd3_n <- scd3_n/max(scd3_n)

####################################################

# Region 4
scd4_n <- as.matrix(stats::dist(select(sc4_n, long, lat), diag = TRUE, upper = TRUE))
scd4_n <- as.data.frame(scd4_n)
colnames(scd4_n) <- sc4_n$State
rownames(scd4_n) <- sc4_n$State
scd4_n <- scd4_n/max(scd4_n)
```




Create new distance matrix for each region for fitness function based on:
\[
    \frac{Patient}{Cost \times Distance \times Hospital Performance Rating} 
\]

```{r}
# Region 1
crp1_n <- sc1_n$crp
crp1_n <- as.data.frame(crp1_n)
row.names(crp1_n) <- sc1_n$State

## transpose
crp1_n <- t(crp1_n)
crp1_n <- as.data.frame(crp1_n)

## Create new distance matrix accounting all variables
scd1_wn <- data.frame(mapply(`*`, scd1_n, crp1_n))
colnames(scd1_wn) <- sc1_n$State
row.names(scd1_wn) <- sc1_n$State

####################################################

# Region 2
crp2_n <- sc2_n$crp
crp2_n <- as.data.frame(crp2_n)
row.names(crp2_n) <- sc2_n$State

## transpose
crp2_n <- t(crp2_n)
crp2_n <- as.data.frame(crp2_n)

## Create new distance matrix accounting all variables
scd2_wn <- data.frame(mapply(`*`, scd2_n, crp2_n))
colnames(scd2_wn) <- sc2_n$State
row.names(scd2_wn) <- sc2_n$State

####################################################

# Region 3
crp3_n <- sc3_n$crp
crp3_n <- as.data.frame(crp3_n)
row.names(crp3_n) <- sc3_n$State

## transpose
crp3_n <- t(crp3_n)
crp3_n <- as.data.frame(crp3_n)

## Create new distance matrix accounting all variables
scd3_wn <- data.frame(mapply(`*`, scd3_n, crp3_n))
colnames(scd3_wn) <- sc3_n$State
row.names(scd3_wn) <- sc3_n$State

####################################################

# Region 4
crp4_n <- sc4_n$crp
crp4_n <- as.data.frame(crp4_n)
row.names(crp4_n) <- sc4_n$State

## transpose
crp4_n <- t(crp4_n)
crp4_n <- as.data.frame(crp4_n)

## Create new distance matrix accounting all variables
scd4_wn <- data.frame(mapply(`*`, scd4_n, crp4_n))
colnames(scd4_wn) <- sc4_n$State
row.names(scd4_wn) <- sc4_n$State
```

Apply the fitness function to GA for each region.

```{r}
# Region 1
nr <- nrow(scd1_wn)
GA.fitr1_n <- ga(
  type = "permutation", 
  fitness = fitness1, 
  distMatrix = scd1_wn, 
  lower = 1, 
  upper = nr, 
  popSize = 100, 
  maxiter = 5000, 
  run = 1000, 
  pmutation = 0.2, 
  monitor = NULL
)

summary(GA.fitr1_n)

####################################################

# Region 2
nr <- nrow(scd2_wn)
GA.fitr2_n <- ga(
  type = "permutation", 
  fitness = fitness1, 
  distMatrix = scd2_wn, 
  lower = 1, 
  upper = nr, 
  popSize = 100, 
  maxiter = 5000, 
  run = 1000, 
  pmutation = 0.2, 
  monitor = NULL
)

summary(GA.fitr2_n)

####################################################

# Region 3
nr <- nrow(scd3_wn)
GA.fitr3_n <- ga(
  type = "permutation", 
  fitness = fitness1, 
  distMatrix = scd3_wn, 
  lower = 1, 
  upper = nr, 
  popSize = 100, 
  maxiter = 5000, 
  run = 1000, 
  pmutation = 0.2, 
  monitor = NULL
)

summary(GA.fitr3_n)

####################################################

# Region 4
nr <- nrow(scd4_wn)
GA.fitr4_n <- ga(
  type = "permutation", 
  fitness = fitness1, 
  distMatrix = scd4_wn, 
  lower = 1, 
  upper = nr, 
  popSize = 100, 
  maxiter = 5000, 
  run = 1000, 
  pmutation = 0.2, 
  monitor = NULL
)

summary(GA.fitr4_n)
```


Prepare data for visualization

```{r}
# Region 1
n = nrow(sc1_n)
id = list(1:n)
id <- as.data.frame(id)
tour <- GA.fitr1_n@solution[1,]
tour <- as.data.frame(tour)
path <- list(1:n)
path <- as.data.frame(path)
tour <- cbind(tour, path)
colnames(tour) <- c("id", "path_order")
vis1 <- cbind(id, sc1_n)
colnames(vis1)[1] <- "id"
vis1 <- merge(vis1, tour)
vis1 <- vis1 %>% 
  arrange(path_order)

####################################################

# Region 2
n = nrow(sc2_n)
id = list(1:n)
id <- as.data.frame(id)
tour <- GA.fitr2_n@solution[1,]
tour <- as.data.frame(tour)
path <- list(1:n)
path <- as.data.frame(path)
tour <- cbind(tour, path)
colnames(tour) <- c("id", "path_order")
vis2 <- cbind(id, sc2_n)
colnames(vis2)[1] <- "id"
vis2 <- merge(vis2, tour)
vis2 <- vis2 %>% 
  arrange(path_order)

####################################################

# Region 3
n = nrow(sc3_n)
id = list(1:n)
id <- as.data.frame(id)
tour <- GA.fitr3_n@solution[1,]
tour <- as.data.frame(tour)
path <- list(1:n)
path <- as.data.frame(path)
tour <- cbind(tour, path)
colnames(tour) <- c("id", "path_order")
vis3 <- cbind(id, sc3_n)
colnames(vis3)[1] <- "id"
vis3 <- merge(vis3, tour)
vis3 <- vis3 %>% 
  arrange(path_order)

####################################################

# Region 4
n = nrow(sc4_n)
id = list(1:n)
id <- as.data.frame(id)
tour <- GA.fitr4_n@solution[1,]
tour <- as.data.frame(tour)
path <- list(1:n)
path <- as.data.frame(path)
tour <- cbind(tour, path)
colnames(tour) <- c("id", "path_order")
vis4 <- cbind(id, sc4_n)
colnames(vis4)[1] <- "id"
vis4 <- merge(vis4, tour)
vis4 <- vis4 %>% 
  arrange(path_order)
```



Visualization


```{r}
p10 <- ggplot() +
  geom_polygon(
    data = US,
    aes(
      x = long,
      y = lat,
      group = group,
      fill = NA
    ),
    color = "grey35",
    fill = NA,
    size = 0.5
  ) + geom_point(
    data = sc_n,
    shape = 13,
    aes(
      x = long,
      y = lat,
      colour = factor(region),
      size = (1/crp),
    )
  ) + geom_path(
  data = vis1, 
  color = "red",
  size = 1.2,
  aes(
    x = long,
    y = lat,
    alpha = 0.9
  )
)+ geom_path(
  data = vis2,
  color = "green",
  size = 1.2,
  aes(
    x = long,
    y = lat,
    alpha = 0.9
  )
)+ geom_path(
  data = vis3, 
  color = "blue",
  size = 1.2,
  aes(
    x = long,
    y = lat,
    alpha = 0.9
  )
)+ geom_path(
  data = vis4, 
  color = "purple",
  size = 1.2,
  aes(
    x = long,
    y = lat,
    alpha = 0.9
  )
) + coord_map() + labs(
  x = NULL,
  y = NULL,
  title = "State Centers - FF2 - Normalized - Kmeans"
) + theme(
  axis.line =  element_blank(),
  axis.text =  element_blank(),
  axis.ticks =  element_blank(),
  legend.position = "none",
  plot.title = element_text(hjust = 0.5, size = 18),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank()
)

p10

# Save image file
pdf(
  file = "image/p10.pdf",
  width = 12,
  height = 7
)

p10

dev.off()

jpeg(
  file = "image/sc_ff2_nor_kmeans.jpeg"
)

p10

dev.off()
```
























